NeuronUnit: Collaboratively Testing the Validity of Neuroscientific Models

### 2013-11-30 NeurounitFrontiers/Paper.tex
Added paths to figure captions

  def _judge(self, model):
    inputs, means, stds = self.inputs, self.means, self.stds
    n = len(inputs)
    counts = numpy.empty((n,))
    for i in xrange(n):
      counts[i] = model.spike_count_from_current(inputs[i])
    chisquared = sum((counts-means)**2 / means) # An array of chi-squared values.  
    p = scipy.stats.chi2.cdf(chisquared,n-1) # An array of p-values.  
    pooled_p = sciunit.utils.fisherp(p_array) # A pooled p-value.  
    return sciunit.PValue(pooled_p, related_data={
      "inputs": inputs, "counts": counts, "obs_means": means, "obs_stds": stds
    })
\end{python}
\vspace{-15px}
\caption{An example single neuron spike count test class implemented using \textit{SciUnit}. Because this implementation contains logic common to many different systems, \textit{NeuronUnit} was developed to provide a simpler means to deliver it (see Sec. \ref{sec:neuronunit}).}
\label{fig:rate_test}
\vspace{-10px}
\end{figure}

@@

\begin{figure}
\begin{python}
class SpikeCountFromCurrent(sciunit.Capability):
  def spike_count_from_current(self, input): 
    """Takes a numpy array containing current stimulus (in nA) and
    produces an integer spike count. Can be called multiple times."""
    raise NotImplementedError("Model does not implement capability.")
\end{python}
\vspace{-15px}
\caption{An example capability specifying a single required method (used by the test in Figure \ref{fig:rate_test}).}
\label{fig:capability}
\vspace{-10px}
\end{figure}
\begin{figure}
\begin{python}
class TrainSpikeCountFromCurrent(sciunit.Capability):
  def train_with_currents(self, currents, counts):
    """Takes a list of numpy arrays containing current stimulus (in nA) and
    observed spike counts. Model parameters should be adjusted based on this
    training data."""
    raise NotImplementedError("Model does not implement capability.")
\end{python}
\vspace{-15px}
\caption{Another capability specifying a training protocol (not used by the test in Figure \ref{fig:rate_test}).}
\label{fig:training}
\vspace{-10px}
\end{figure}

@@

  def train_with_currents(self, currents, counts):
    means = [numpy.mean(c) for c in currents]
    [self.offset, self.scale] = numpy.polyfit(means, counts, deg=1)    
\end{python}
\vspace{-15px}
\caption{A model that returns a spike count by applying a linear transformation to the mean input current. The parameters can be provided manually or learned from data provided by a test or user (see text).}
\label{fig:simple_model}
\vspace{-10px}
\end{figure}

@@

\subsection{Test Suites and Score Matrices} A collection of tests intended to be run on the same model can be put together to form a test suite.
The following is a test suite that could be used for a simplified version of the QSNMC, as described above:  
\begin{python}
CA1_suite = sciunit.TestSuite([CA1_sc_test, CA1_sl_test, CA1_sv_test, CA1_isi_test, CA1_ism_test])
\end{python}
Like a single test, a test suite is capable of judging one or more models. The result is a score matrix much like the one diagramed in Fig. \ref{fig:sciunit_overview}.
\begin{python}
scores = CA1_suite.judge([CA1_linear_model_heuristic, CA1_linear_model_fit])
\end{python}
A simple summary of the scores in a score matrix can be printed to the console or visualized by other tools in development.  For example, the score matrix might be visualized as a hyperlinked table inside an IPython notebook containing the specification and execution of the test suite (Fig. \ref{fig:scidash_matrix}). Such notebooks are stored in test repositories like \textit{hippocampus\_tests}.

@@

\subsection{A Collaborative Workflow}\label{sec:neuronunit_acitivities}
The \textit{NeuronUnit} package contains a collection of test families and associated capabilities relevant to neurophysiology, such as the tests used in the QSNMC, and is collaboratively developed on GitHub (\url{http://github.com/scidash/neuronunit}). 
Scientists can use these tools to easily create test suites tailored to specific neuronal systems of interest. 
For example, scientists interested specifically in testing models of cells in the hippocampus would use \textit{NeuronUnit} to create tests in a GitHub repository called \textit{hippocampus\_tests}, parameterizing them by relevant data from \textit{NeuroElectro}. 
To make this easy, \textit{NeuronUnit} contains \emph{bridge functions} for retrieving data and generating tests from \textit{NeuroElectro} directly (Sec. \ref{sec:neuroelectro}).

Statisticians interested in improving how goodness-of-fit is measured can simply fork the \textit{NeuronUnit} repository, modify the logic in the relevant test families and submit a \emph{pull request} for evaluation by the community. 
Rather than attempting to convince a large number of scientists socially that the methods used in canonical papers need to be changed, statisticians need only propagate a change to the testing logic used by the community. 
Once the testing logic has been changed, all models, past and future, will be immediately re-evaluated against the new metrics. 
Scientists wanting to determine statistical best-practices for measuring goodness-of-fit need only refer to the tests in these repositories.

>>>>

  def _judge(self, model):
    inputs, means, stds = self.inputs, self.means, self.stds
    n = len(inputs)
    counts = numpy.empty((n,))
    for i in xrange(n):
      counts[i] = model.spike_count_from_current(inputs[i])
    chisquared = sum((counts-means)**2 / means) # An array of chi-squared values.  
    p = scipy.stats.chi2.cdf(chisquared,n-1) # An array of p-values.  
    pooled_p = sciunit.utils.fisherp(p_array) # A pooled p-value.  
    return sciunit.PValue(pooled_p, related_data={
      "inputs": inputs, "counts": counts, "obs_means": means, "obs_stds": stds
    })
\end{python}
\vspace{-15px}
\caption{\texttt{[neuronunit.SpikeCountTest]} An example single neuron spike count test class implemented using \textit{SciUnit}. Because this implementation contains logic common to many different systems, \textit{NeuronUnit} was developed to provide a simpler means to deliver it (see Sec. \ref{sec:neuronunit}).}
\label{fig:rate_test}
\vspace{-10px}
\end{figure}

@@

\begin{figure}
\begin{python}
class SpikeCountFromCurrent(sciunit.Capability):
  def spike_count_from_current(self, input): 
    """Takes a numpy array containing current stimulus (in nA) and
    produces an integer spike count. Can be called multiple times."""
    raise NotImplementedError("Model does not implement capability.")
\end{python}
\vspace{-15px}
\caption{\texttt{[neuronunit.SpikeCountFromCurrent]} An example capability specifying a single required method (used by the test in Figure \ref{fig:rate_test}).}
\label{fig:capability}
\vspace{-10px}
\end{figure}
\begin{figure}
\begin{python}
class TrainSpikeCountFromCurrent(sciunit.Capability):
  def train_with_currents(self, currents, counts):
    """Takes a list of numpy arrays containing current stimulus (in nA) and
    observed spike counts. Model parameters should be adjusted based on this
    training data."""
    raise NotImplementedError("Model does not implement capability.")
\end{python}
\vspace{-15px}
\caption{\texttt{[neuronunit.TrainSpikeCountFromCurrent]} Another capability specifying a training protocol (not used by the test in Figure \ref{fig:rate_test}).}
\label{fig:training}
\vspace{-10px}
\end{figure}

@@

  def train_with_currents(self, currents, counts):
    means = [numpy.mean(c) for c in currents]
    [self.offset, self.scale] = numpy.polyfit(means, counts, deg=1)    
\end{python}
\vspace{-15px}
\caption{\texttt{[neuronunit.LinearModel]} A model that returns a spike count by applying a linear transformation to the mean input current. The parameters can be provided manually or learned from data provided by a test or user (see text).}
\label{fig:simple_model}
\vspace{-10px}
\end{figure}

@@

\subsection{Test Suites and Score Matrices} A collection of tests intended to be run on the same model can be put together to form a test suite.
The following is a test suite that could be used for a simplified version of the QSNMC, as described above:  
\begin{python}
CA1_suite = sciunit.TestSuite([CA1_sc_test, CA1_sl_test, CA1_sv_test, CA1_isi_test, CA1_ism_test])
\end{python}
Like a single test, a test suite is capable of judging one or more models. The result is a score matrix much like the one diagramed in Fig. \ref{fig:sciunit_overview}.
\begin{python}
scores = CA1_suite.judge([CA1_linear_model_heuristic, CA1_linear_model_fit])
\end{python}
A simple summary of the scores in a score matrix can be printed to the console or visualized by other tools in development.  For example, the score matrix might be visualized as a hyperlinked table inside an IPython notebook \todo{reference to ipython nb docs/paper?} containing the specification and execution of the test suite (Fig. \ref{fig:scidash_matrix}). Such notebooks are stored in test repositories like \textit{hippocampus\_tests}.

@@

\subsection{A Collaborative Workflow}\label{sec:neuronunit_acitivities}
The \textit{NeuronUnit} package contains a collection of test families and associated capabilities relevant to neurophysiology, such as the tests used in the QSNMC, and is collaboratively developed on GitHub (\url{http://github.com/scidash/neuronunit}). 
Scientists can use this package to easily create test suites tailored to specific neuronal systems of interest. 
For example, scientists interested specifically in testing models of cells in the hippocampus would use \textit{NeuronUnit} to create tests in a GitHub repository containing an iPython notebook, as described above, called \textit{hippocampus\_tests}, parameterizing them by relevant data from \textit{NeuroElectro}. 
To make this easy, \textit{NeuronUnit} contains \emph{bridge functions} for retrieving data and generating tests from \textit{NeuroElectro} directly (Sec. \ref{sec:neuroelectro}).

Statisticians interested in improving how goodness-of-fit is measured can simply fork the \textit{NeuronUnit} repository, modify the logic in the relevant test families and submit a \emph{pull request} for evaluation by the community. 
Rather than attempting to convince a large number of scientists socially that the methods used in canonical papers need to be changed, statisticians need only propagate a change to the testing logic used by the community. 
Once the testing logic has been changed, all past models will be immediately evaluated against the new metrics. 
Scientists wishing to determine statistical best-practices for measuring goodness-of-fit need only refer to the tests in these repositories.
